A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-08-14 code/supplemental_v2.Rmd
Update supplemental s2 with key, inprogress s4 -does not run using loop

cancer_w_subtypes <- list()
for(n in cancer_type) {
  spec_cancer <- umap_cancertype %>% 
    mutate(J_color = case_when(project_code == n ~ Subtype_Selected,
                               TRUE ~ "Other")) %>%
    filter(J_color != "Other")
  if(nrow(spec_cancer) == 0) {next
  }
  cancer_w_subtypes[[n]] <- n
}
cancer_w_subtypes <- names(cancer_w_subtypes)

@@

  with(spec_cancer,
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = J_colors[J_color], pch = 21, cex = 0.8, lwd = 0.2,
               theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
  legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
#legend("top", inset = c(0), legend=names(J_colors[1:2]), col=J_colors[1:2], pch = 20, bty = "n", horiz = T, cex = 1.2)
#legend("top", inset = c(0.10), legend=names(J_colors[3:4]), col=J_colors[3:4], pch = 20, bty = "n", horiz = T, cex = 1.2)
  title(xlab="UMAP 2", line=0)
}

@@

### Supplementary Figure S4
This is the 377 pathway UMAP. I will need to develop a loop that does this.

**Supplementary Figure S4. Projection of pathway disruptions onto the UMAP.** Each panel shows one of 377 Reactome
pathways projected on the UMAP from **Figure 1D of the main text** ordered by the hierarchical pathway order 
(column) from **Figure 2C of the main text**.
```{r supplementary-figure-s4-data-generation, eval = FALSE}
load("../data/gdc_reactome_path_analysis.Rda") ## load data
umap_cancertype <- fread("../data/umap_3d_coors.tsv") 
n_path <- 377 ## how many pathways do we want to look at starting from the most to least mutated pathway

@@

selected_path <- umap_cancertype %>% dplyr::select(sample_id, plot_x, plot_y) %>% ## select relevant columns
  left_join(., selected_path, by = "sample_id") %>%
  melt(data = ., id.vars = c("sample_id", "plot_x", "plot_y")) %>%
  filter(value != 0)

@@

load("../data/selected_path.Rda") ## load data
load("../data/gdc_reactome_path_analysis.Rda")
umap_cancertype <- fread("../data/umap_3d_coors.tsv")
umap_bg <- umap_cancertype[,-"project_code"] ## get background data for facet

n_path <- 377 ## number of pathways we want to see

toppath <- gdc_reactome_path_analysis %>% rowSums() %>% as.data.frame() %>% rownames_to_column(var = "path_name") %>% arrange(desc(.)) %>% .[1:n_path,] %>% .[, "path_name"] ## order pathway by most to least mutated pathway.. needed for facet

## creating layout to plot all 377 pathways
figs4_layout <- data.frame(start = seq(from = 1, to = length(toppath), by = 25),
                          end = c(seq(from = 25, to = length(toppath), by = 25), length(toppath)))

for(n in 1:nrow(figs4_layout)) {
  message(n)
  pathway_layout <- figs4_layout %>% .[n,] ## getting the layout for selecting pathways
  pathway_layout <- toppath[pathway_layout$start:pathway_layout$end] ## filtering those pathways from toppath
  filtered_path <- selected_path %>% filter(variable %in% pathway_layout) %>% ## filtering our data to only include those pathways
    mutate(variable = factor(variable, levels = c(pathway_layout)))
  ggplot_umap <- ggplot(filtered_path, aes(x = plot_y, y = plot_x, colour = as.character(value))) + 
    geom_point(data = umap_bg, aes(x = plot_y, y = plot_x), colour = "grey", size = 0.25) +
    geom_point(size = 0.25) +
    theme_classic() +
    coord_fixed() +
    xlab("UMAP 1") + ylab("UMAP 2") +
    scale_colour_manual(values = rep("firebrick3", 25)) +
    theme(axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks = element_blank(),
      axis.line = element_blank(),
      strip.text = element_text(size = 5),
      legend.position = "none") +
    facet_wrap(~ variable,
               ncol = 5) ## facet by pathways
  new_dfname <- paste0("ggplot_umap_", n)
  assign(paste0(new_dfname), value = ggplot_umap)
}

save(ggplot_umap_1,ggplot_umap_2,ggplot_umap_3,ggplot_umap_4,ggplot_umap_5,ggplot_umap_6,ggplot_umap_7,ggplot_umap_8,ggplot_umap_9,ggplot_umap_10,ggplot_umap_11,ggplot_umap_12,ggplot_umap_13,ggplot_umap_14,ggplot_umap_15,ggplot_umap_16, file = "../data/fig_s4_ggplot.Rda")
```

>>>>

cancer_w_subtypes <- list()
for(n in cancer_type) {
  spec_cancer <- umap_cancertype %>% 
    mutate(J_color = case_when(project_code == n ~ Subtype_Selected,
                               TRUE ~ "Other")) %>%
    filter(J_color != "Other")
  if(nrow(spec_cancer) == 0) {next
  }
  cancer_w_subtypes[[n]] <- n
}

cancer_w_subtypes <- names(cancer_w_subtypes)

@@

  with(spec_cancer,
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = J_colors[J_color], pch = 21, cex = 0.8, lwd = 0.2,
               theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
  legend("top", inset = c(0.75), legend=n, bty = "n", horiz = T, cex = 1.2)
  legend("top", inset = c(0), legend=names(J_colors[1:2]), col=J_colors[1:2], pch = 20, bty = "n", horiz = T, cex = 1.2)
  legend("top", inset = c(0.10), legend=names(J_colors[3:4]), col=J_colors[3:4], pch = 20, bty = "n", horiz = T, cex = 1.2)
  legend("top", inset = c(0.20), legend=names(J_colors[5:6]), col=J_colors[5:6], pch = 20, bty = "n", horiz = T, cex = 1.2)
  title(xlab="UMAP 2", line=0)
}

@@

```{r supplementary-figure-s4-data-generation, eval = FALSE}
load("../data/gdc_reactome_path_analysis.Rda") ## load data
umap_cancertype <- fread("../data/umap_3d_coors.tsv") 
n_path <- 377 ## how many pathways do we want to look at starting from the most to least mutated pathway

@@

selected_path <- umap_cancertype %>% dplyr::select(sample_id, plot_x, plot_y, plot_z) %>% ## select relevant columns
  left_join(., selected_path, by = "sample_id") %>%
  melt(data = ., id.vars = c("sample_id", "plot_x", "plot_y", "plot_z")) %>%
  filter(value != 0)

@@

# load("../data/selected_path.Rda") ## load data
# load("../data/gdc_reactome_path_analysis.Rda")
# umap_cancertype <- fread("../data/umap_3d_coors.tsv")
# umap_bg <- umap_cancertype[,-"project_code"] ## get background data for facet
# 
# n_path <- 377 ## number of pathways we want to see
# 
# toppath <- gdc_reactome_path_analysis %>% rowSums() %>% as.data.frame() %>% rownames_to_column(var = "path_name") %>% arrange(desc(.)) %>% .[1:n_path,] %>% .[, "path_name"] ## order pathway by most to least mutated pathway.. needed for facet
# 
# ## creating layout to plot all 377 pathways
# figs4_layout <- data.frame(start = seq(from = 1, to = length(toppath), by = 25),
#                           end = c(seq(from = 25, to = length(toppath), by = 25), length(toppath)))
# 
# for(n in 1:nrow(figs4_layout)) {
#   message(n)
#   pathway_layout <- figs4_layout %>% .[n,] ## getting the layout for selecting pathways
#   pathway_layout <- toppath[pathway_layout$start:pathway_layout$end] ## filtering those pathways from toppath
#   filtered_path <- selected_path %>% filter(variable %in% pathway_layout) %>% ## filtering our data to only include those pathways
#     mutate(variable = factor(variable, levels = c(pathway_layout)))
#   ggplot_umap <- ggplot(filtered_path, aes(x = plot_y, y = plot_x, colour = as.character(value))) + 
#     geom_point(data = umap_bg, aes(x = plot_y, y = plot_x), colour = "grey", size = 0.25) +
#     geom_point(size = 0.25) +
#     theme_classic() +
#     coord_fixed() +
#     xlab("UMAP 1") + ylab("UMAP 2") +
#     scale_colour_manual(values = rep("firebrick3", 25)) +
#     theme(axis.text.x = element_blank(),
#       axis.text.y = element_blank(),
#       axis.title.x = element_blank(),
#       axis.title.y = element_blank(),
#       axis.ticks = element_blank(),
#       axis.line = element_blank(),
#       strip.text = element_text(size = 5),
#       legend.position = "none") +
#     facet_wrap(~ variable,
#                ncol = 5) ## facet by pathways
#   new_dfname <- paste0("ggplot_umap_", n)
#   assign(paste0(new_dfname), value = ggplot_umap)
# }
# 
# save(ggplot_umap_1,ggplot_umap_2,ggplot_umap_3,ggplot_umap_4,ggplot_umap_5,ggplot_umap_6,ggplot_umap_7,ggplot_umap_8,ggplot_umap_9,ggplot_umap_10,ggplot_umap_11,ggplot_umap_12,ggplot_umap_13,ggplot_umap_14,ggplot_umap_15,ggplot_umap_16, file = "../data/fig_s4_ggplot.Rda")
```

```{r figure-s4-plot, fig.width= 9}
load("../data/selected_path.Rda")
load("../data/effect_heatmap.Rda") ## load effect heatmap data
pathorder <- as.character(levels(effect_heatmap$path_name)) ## getting col order

startingpoint <- seq(1, 377, by = 15)
startingpointlist <- list()
for(n in startingpoint) {
  start1 <- n
  end1 <- start1 + 5
  start2 <- end1 + 1
  end2 <- start2 + 5
  start3 <- end2 + 1
  end3 <- start3 + 2
  startingpointlist[[n]] <- data.frame(start1 = start1,
                                       end1 = end1,
                                       start2 = start2,
                                       end2 = end2,
                                       start3 = start3,
                                       end3 = end3)
}
startingpointlist <- as.data.frame(do.call(rbind, startingpointlist))

s4plot <- function(x) {
  a <-  x[,"start1"]
  b <- x[,"end1"]
  c <- x[,"start2"]
  d <- x[,"end2"]
  e <- x[,"start3"]
  f <- x[,"end3"]
par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in pathorder[x[,"start1"]:x[,"end1"]]) {
  pathname <- as.data.frame(strsplit(n, " ")) ## split string into word
  pathname <- pathname[,1] ## get the character vectors
  with(selected_path %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
path1 <- paste0(pathname[1], " ", pathname[2], " ", pathname[3], " ", pathname[4], " ")
path1 <- gsub("NA", "", path1)
path2 <- paste0(pathname[5], " ", pathname[6], " ", pathname[7], " ", pathname[8], " ")
path2 <- gsub("NA", "", path2)
legend("top", inset = c(0.01), legend=path1, bty = "n", horiz = T, cex = 0.9)
legend("top", inset = c(0.06), legend=path2, bty = "n", horiz = T, cex = 0.9)
}

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in pathorder[x[,"start2"]:x[,"end2"]]) {
  pathname <- as.data.frame(strsplit(n, " ")) ## split string into word
  pathname <- pathname[,1] ## get the character vectors
  with(selected_path %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
path1 <- paste0(pathname[1], " ", pathname[2], " ", pathname[3], " ", pathname[4], " ")
path1 <- gsub("NA", "", path1)
path2 <- paste0(pathname[5], " ", pathname[6], " ", pathname[7], " ", pathname[8], " ")
path2 <- gsub("NA", "", path2)
legend("top", inset = c(0.01), legend=path1, bty = "n", horiz = T, cex = 0.9)
legend("top", inset = c(0.06), legend=path2, bty = "n", horiz = T, cex = 0.9)
}

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in pathorder[x[,"start3"]:x[,"end3"]]) {
  pathname <- as.data.frame(strsplit(n, " ")) ## split string into word
  pathname <- pathname[,1] ## get the character vectors
  with(selected_path %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
path1 <- paste0(pathname[1], " ", pathname[2], " ", pathname[3], " ", pathname[4], " ")
path1 <- gsub("NA", "", path1)
path2 <- paste0(pathname[5], " ", pathname[6], " ", pathname[7], " ", pathname[8], " ")
path2 <- gsub("NA", "", path2)
legend("top", inset = c(0.01), legend=path1, bty = "n", horiz = T, cex = 0.9)
legend("top", inset = c(0.06), legend=path2, bty = "n", horiz = T, cex = 0.9)
}
}

a = 1
b = 6
c = 7
d = 12
e = 13
f = 15

for(p in 1:nrow(startingpointlist)-1) {
  data <- startingpointlist[p,]
  s4plot(x = data)
}

# par(mfrow = c(2,3), mai = c(0.1, 0.1, 2, 0.1))
# par(mar = c(2,2,0,0))
# for(n in genes[7:12]) {
# with(selected_mut %>% 
#        filter(variable == n),
#      scatter3D(plot_y, -plot_z, plot_x, 
#                                 bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
#                                 theta = 0, phi = 65, scale = F,
#                xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
#                                 xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
#                                 colvar = NULL))
# title(xlab="UMAP 2", line=0)
# legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
# }
# par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
# par(mar = c(2,2,0,0))
# for(n in genes[13:15]) {
# with(selected_mut %>% 
#        filter(variable == n),
#      scatter3D(plot_y, -plot_z, plot_x, 
#                                 bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
#                                 theta = 0, phi = 65, scale = F,
#                xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
#                                 xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
#                                 colvar = NULL))
# title(xlab="UMAP 2", line=0)
# legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
# }

```
**Supplementary Figure S4. Projection of pathway disruptions onto the UMAP.** Each panel shows one of 377 Reactome
pathways projected on the UMAP from **Figure 1D of the main text** ordered by the hierarchical pathway order 
(column) from **Figure 2C of the main text**.

### 2020-08-14 code/survival/survfig.Rmd
Update supplemental s2 with key, inprogress s4 -does not run using loop

```{r, survfig, echo=FALSE, warning=FALSE, error=TRUE,message=FALSE,fig.height=9, fig.width=6.5}

@@

mylayout = '
AAAAAAAA
BBBBBBBB
BBBBBBBB
CCCCCCCC
CCCCCCCC
CCCCCCCC
'

print(
((p1+mytheme)/((pstad[[1]]+mytheme)+(pgbm[[1]]+mytheme) +
(ppaad[[1]]+mytheme) + (thca[[1]] + mytheme)))/((pcanc+mytheme) + (pclass+mytheme))+plot_layout(design=mylayout) +plot_annotation(tag_levels = "A")
)

```

>>>>

#### Figure 7
```{r, survfig, echo=FALSE, warning=FALSE, error=TRUE,message=FALSE,fig.height=9, fig.width=9}

@@

kmlayout = "
AB
AB
AB
CD
CD
CD
EE
"
kmlegend <- theme(legend.direction = "horizontal")

print(
((p1+mytheme)/((pstad[[1]]+mytheme + kmlegend)+(pgbm[[1]]+mytheme + kmlegend) +
(ppaad[[1]]+mytheme + kmlegend) + (thca[[1]] + mytheme + kmlegend) + guide_area() + plot_layout(design = kmlayout, guides = "collect"))) + plot_annotation(tag_levels = "A")
)
```

```{r, results='asis'}
cat("\\newpage")
```

#### Figure 8
```{r, message = FALSE, fig.width = 9}
mylayout = '
ABB
ABB
#BB
'

((pcanc+mytheme) + (pclass+mytheme))+plot_layout(design=mylayout) +plot_annotation(tag_levels = "A")
```

### 2020-08-14 text/manuscript.Rmd
Update supplemental s2 with key, inprogress s4 -does not run using loop

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 1
```{r child='../code/figure1v2.Rmd'}
#
```

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 2
```{r child='../code/figure2v2.Rmd'}
#
```

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 3
```{r child = '../code/figure4v2.Rmd'}
#
```

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 4
```{r child = '../code/figure3v2.Rmd'}
#
```

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 5
```{r child = '../code/figure2.Rmd'}
#
```

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 6
```{r child = '../code/figure4.Rmd'}
#
```
#### Figure 7
```{r, results='asis'}
cat("\\newpage")
```

@@

```{r, results='asis'}
cat("\\newpage")
```
#### Table 1
```{r child = '../code/table2.Rmd'}
#
```

>>>>

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 1
```{r }
#child='../code/figure1v2.Rmd'
```

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 2
```{r }
#child='../code/figure2v2.Rmd'
```

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 3
```{r }
#child = '../code/figure4v2.Rmd'
```

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 4
```{r }
#child = '../code/figure3v2.Rmd'
```

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 5
```{r }
#child = '../code/figure2.Rmd'
```

```{r, results='asis'}
cat("\\newpage")
```
#### Figure 6
```{r }
#child = '../code/figure4.Rmd'
```

@@

```{r, results='asis'}
cat("\\newpage")
```
#### Table 1
```{r }
#child = '../code/table2.Rmd'
```
